A Constraint Solving Approach to Parikh Images of Regular Languages

### 2023-07-11 generalised.tex
comments, bit of shortening
more

Observe that every homomorphism~$\Map:\: \Sigma^* \to \Monoid$ can be
represented as the composition~$h' \circ \ParikhMap$, for some
homomorphism~$h' : \Naturals^k \to M$. One of the insights underlying our
approach is that it is more efficient to directly compute a
projection on the Parikh image~$\Map(\Language)$, than to first compute the
standard image~$\ParikhMap(\Language)$ followed by projection to
some property of interest.

@@

% @Philipp: I am quite unsure about which intersection we are looking for here,
% please verify I'm not wrong!
% PR: Sounds right (did some small fixes)
Denoting the language described by \cref{fig:aut_a} as $\Language_{pre}$, we can
then replace $s_1 = \Substring(s_2, i, n)$ and $s_2 \in
\Language(B)$ (of \cref{const:s1-in-c-dd}) with an
equi-satisfiable formula that no longer contains any explicit substring
operation:
%
\begin{equation}
  \label{eq:parikh-constraint}
  p \in \Language_{pre} \wedge~
  \pi_\Alphabet(p) \in \Language(B)
  \wedge~~ \pi_D(p) =
  \begin{bmatrix}
    i \\ n
  \end{bmatrix}~~
  \wedge 0 \leq i \leq i + n \leq |p|
\end{equation}
%
To check the satisfiability of \cref{eq:parikh-constraint}, we need a
decision procedure that can process intersections of regular languages
(in this case, of $\Language_{pre}$ and $\Language(B)$, synchronising
on $\Alphabet$), while imposing the side
condition~$0 \leq n \leq m \leq |p|$ on the increment sum. In
\cite{ostrich-plus}, this decision procedure turned out to be main
bottleneck of the string solver, which was one of the motivations to
develop the lazy algorithm proposed in this paper.

>>>>

Observe that every homomorphism~$\Map:\: \Sigma^* \to \Monoid$ can be
represented as the composition~$h' \circ \ParikhMap$, for some
homomorphism~$h' : \Naturals^k \to M$. One of the insights underlying our
approach is that it is more efficient to directly compute a
projection~$\Map(\Language)$ on the Parikh image, than to first compute the
standard image~$\ParikhMap(\Language)$ followed by projection to
some property of interest.

@@

% AS: I am quite unsure about which intersection we are looking for here,
% please verify I'm not wrong!
% PR: Sounds right (did some small fixes)
Denoting the language described by \cref{fig:aut_a} as $\Language_{pre}$, we can
then replace $s_1 = \Substring(s_2, i, n)$ and $s_2 \in
\Language(B)$ (of \cref{const:s1-in-c-dd}) with an
equi-satisfiable formula that no longer contains any explicit substring
operation:
%
\begin{equation}
  \label{eq:parikh-constraint}
  p \in \Language_{pre} \wedge~
  \pi_\Alphabet(p) \in \Language(B)
  \wedge~~ \pi_D(p) =
  \begin{bmatrix}
    i \\ n
  \end{bmatrix}~~
  \wedge 0 \leq i \leq i + n \leq |p|
\end{equation}
%
To check the satisfiability of \cref{eq:parikh-constraint}, we need a
decision procedure that can process intersections of regular languages
(in this case, of $\Language_{pre}$ and $\Language(B)$, synchronising
on $\Alphabet$), while imposing the side
condition~$0 \leq n \leq m \leq |p|$ on the increment sum. In
\cite{ostrich-plus}, this decision procedure turned out to be main
bottleneck of the string solver, which was one of the motivations to
develop the lazy algorithm proposed in this paper.
